---
title: Regression Models Course Project - Motor Trend Data = 'mtcars' Miles Per
  Gallon Analysis
author: "james c walmsley"
date: "12/1/2016"
output:
  pdf_document:
    fig_caption: yes
    fig_height: 5.5
  html_document: default
---
# Executive Summary:                      
        
Using linear regression model variations, including the step function, to gauge model fit,  
we identified model fa <- lm(mpg ~ (wt + qsec + am), data= mtcars) that provides an 84.966  
R^2 value indicating a reasonalby good fit and another model fit fb <- lm(mpg ~ I(cyl + disp  
+ hp + drat + wt + factor(am), data = mtcars) devloped using the multivariate nested approach  
followed by and anova table test to check for multicollinearity which also produced a reasonably
good fit with an R^2 value of 85.13%. It must also be noted that vehicle weight is highly  
correlated (-86.77%) with mpg ratings and transmission type is relatively highly correlated  
with vehicle wieght at (69%).
        
Problem Statement:              
        Backround information, problem statement & questions of interest:
        Background situation:
                As a member of a team of data analysts for the Motor Trend Magazine we have been given  
                a data set called "mtcars" and asked to answer some questions of interest concerning  
                differences between automatic and  manual transmissoin types in regards to associated  
                mpg or miles per gallon ratings in this data set. 
        Assumptions:
                The given data set (a sample of a unknown larger population) for this analysis consists  
                of (iid) independent and identically distrubuted random varialbles for 32 subjects  
                (vehicles) of 11 observations or variables.
                
                Questions of interest for Motor Trend Magazine:
                
                Q1 “Is an automatic or manual transmission better for 'mpg'”
                        or which type of transmission is associated with better mpg or gas mileage ratings?
                                
                        A1. The mean "mpg" rating of all vehicle models including both transmission types  
                        is 20.09 mpg with a 95% confidence interval of 17.92 mpg to 22.26 mpg.
                                
                Q2 "Quantify the mpg difference between automatic and manual transmissions"
                        What is the expected difference in mpg rating and how accurate  
                        is this estimate based on the given data?
                                
                        A2. The mean "mpg" of models with automatic transmisions is 17.15 mpg, with a 95%  
                        confidence interval of between 14.85 mpg to 19.44 mpg and vehicles with manual  
                        transmisions have a mean of 24.39 mpg for a difference of 7.24 mpg with a 95 %  
                        confidence interval of between 18.49 mpg and 30.29 mpg  
                                
Analysis Considerations:                 
        Descriptive - any(is.na()), str() & summary(),
        Exploratory - pairsPlots(), histograms(), boxPlots(), barPlots() QQ_Plots & multiple plots
        Regression Models Analysis -
                SLR, BiVariate Regression, Multivariate Linear Regression, model selection, 
                adjustments, residuals (predict fit, residual fit (-1)), coeficients, correlation,  
                confidence intervals, influence & leverage,
        Diagnostics
                See section on Diagnostics
        Final Model Selection strategy
                Beginning with the simmple linear regression using just one predictor (am),
                Use a bivariate model, Use a multivariate model, Use an intercept adjusted multivariate model
                Use a multivariate model removing a suspected key regressor, Use the nested multivariate process
                Use the step(function) both directions process, Usee different combinations and leave out from  
                the results of the step(function), Choose the best fit which is understandable and easy to explian
        
# Technical Environment:                  
        System - session Info; Set the Working Directory; Record the System & Session Info; Check for required packages

```{r, SetDirectory, echo=TRUE, eval=TRUE, include=FALSE, results='hide'}
setwd("~/Desktop/Coursera_R/7_Regression Models/RM_proj_MPG_MotorTrendData");sessionInfo();installed.packages();date()
```

# Raw Data:                               
        Clean up work space, import the data & check for missing values
        Overview: Motor Trend 'mtcars' data set:
```{r, Data_mtcars, include=FALSE, results='hide'}
rm(list=ls());library(car);library(dplyr);library(MASS);data("mtcars");any(is.na(mtcars))
```
A data frame with 32 observations on 11 variables.

[, 1]	mpg	Miles/(US) gallon
[, 2]	cyl	Number of cylinders (4,6,8)
[, 3]	disp	Displacement (cu.in.)
[, 4]	hp	Gross horsepower
[, 5]	drat	Rear axle ratio
[, 6]	wt	Weight (1000 lbs)
[, 7]	qsec	1/4 mile time
[, 8]	vs	V/S (0 = vee-block, 1 = straight-block)
[, 9]	am	Transmission (0 = automatic, 1 = manual)
[,10]	gear	Number of forward gears (3:5)
[,11]	carb	Number of carburetors (1:4,6,8)

# Processed Data:                         
        Factor columns 2 & 8:11 (cyl,vs,am,gear,carb) so there values can be used as levels
```{r, ProcessData, echo=TRUE, eval=FALSE, cache=TRUE, message=FALSE, warning=FALSE, include=FALSE, results='hide'}
data("mtcars")
for(i in c(2,8:11))mtcars[,i] <- as.factor(mtcars[,i]);
str(mtcars)
```

# Descriptive Statistics:                 

```{r, DescriptiveStats, echo=TRUE, eval=FALSE, cache=TRUE, message=FALSE, warning=FALSE, include=FALSE, results='hide'}
library(datasets);library(dplyr);data("mtcars")
head(mtcars,4);mean(mtcars$mpg);sd(mtcars$mpg)
round(t.test(mtcars$mpg)$conf.int,3)
mtcars0 <- mtcars[mtcars$am==0,];mtcars0;t.test(mtcars0$mpg)
mtcars1 <- mtcars[mtcars$am==1,];mtcars1;t.test(mtcars1$mpg)
c6 <- mtcars$mpg[mtcars$cyl==6];c6
c4 <- mtcars$mpg[mtcars$cyl==4];c4;t.test(c4,c6,var.equal = TRUE)
c6 <- mtcars$mpg[mtcars$cyl==6];c6
c8 <- mtcars$mpg[mtcars$cyl==8];c8;t.test(c6,c8,var.equal = TRUE)
```

# Exploratory Analysis:                   
        See Appendix A. Figures (pairs-plot, histogram, box-plot)
        
# Statistical Modeling:                   
        Multivarite Linear Model Finding Best Fit with Step function:
```{r ModelFitStep, echo=TRUE, eval=FALSE, message=FALSE, warning=FALSE, cache=TRUE, include=TRUE,  results='hide'}
library(stats);library(MASS)
fstp <- lm(mpg ~ ., data = mtcars)
stp <- step(fstp, trace = FALSE)
table(coef(summary(stp)))
summary(stp)$r.squared
```

# Preliminary findings: 
        # Quesions of interest: & interpretation of results:
        A Revisit the Question - Considering all regressors:
                A. Is an automatic or manual transmission better for mpg
                        The results of using multiple linear regression techniques  
                        sugggest that manual transmissions are associated with better 
                        mpg ratings than automatic transmissions
                B. Quantify the MPG difference between automatic and manual transmissions
                        On average manual transmissions provides 24.39 mpg which is 7.24 mpg
                        more than the 17.15 mpg average of the automatic transmission models
        B Primary result
                A. Are any other regressors significantly correlated with mpg rating?
                        a. model fnm6 = factor(am) + cyl + disp + hp + drat + wt
                        this model has an R^2 value of 85.13% 
                B. Further testing
                        a. using the step function in both directions selects wt, qsec and am 
                        as good predictors with an 84.96% R^2 value indicating very good 
                        predictability using this set of regresssors
                
        C Direction, Magnitude, Uncertainty
                A.
#       Multivarite Linear Model Finding Best Fit with Step function:
```{r MVLM_Step_R^2, eval=TRUE, echo=TRUE, fig.align='center', fig.height=4, fig.width=7, message=FALSE, warning=FALSE, include=TRUE, results='asis'}
library(stats);library(MASS);library(ggplot2)
fstp <- lm(mpg ~ ., data = mtcars)
stp <- step(fstp, trace = FALSE)
coef(summary(stp))
summary(stp)$r.squared
par(mfrow = c(1, 1), mar = c(4,4,4,2))
g <- ggplot(mtcars, aes(x = (wt + qsec + am), y = mpg),)
g <- g + xlab("Model stp = (wt + qsec + am)")
g <- g + ylab("Miles per Gallon")
g <- g + geom_smooth(aes(method = "lm", shape = factor(am)))
g <- g + geom_point(aes(shape = factor(am), size=qsec, colour=wt))
g
```
        D Context
                A. It should be noted that vehicle weight has a strong negative correlation
                to mpg ratings (-86.76%) and the weight of vehicle models with manual transmissions  
                range from 1.513tons to 3.570 tons and the weight of vehicles with automatic  
                transmissions range from 2.465 tons to 5.424 tons 
                
        E Implications - Congruence with existing knowledge?
                A. Sedan, Sports, Luxury
                Generally accepted expectations of mpg ratings are that sports and luxury models  
                typically will have lower mpg ratings than sedans


Diagnostics:                             
        Diagnostic tests were conducted on model results in accordance with the plan for analysis  
        considerations. Besides several vehicle models exhibiting leverage on the model fit at the
        high end of the qsec, and weight scales results were generally as expected with faster and  
        or heavier vehilces of both manual and automatic transmissoin types getting lower mpg  
        ratings and slower lighter vehicle models of both transmission types exhibiting better  
        or higher mpg ratings.

Hypothesis Test:                        ?
                H0 = mean(automatic transmission)mpg  =   mean(manual transmission)mpg
                Ha = mean(automatic transmission)mpg  !=  mean(manual transmission)mpg


# Inference & Prediction:                 ?







# Interpretation of Results:              #







# Appendix A. Figures:                    
need to review again possibly condense to one plot













