---
title: "Regression Models Project - Motor Trend Data 'mtcars' Miles Per Gallon Analysis"
author: "james c walmsley"
date: "12/1/2016"
output: pdf_document
---
# EXECUTIVE SUMMARY
        Information to be added!





# Problem Statement / Define the Question(s)
        You work for Motor Trend, a magazine about the automobile industry. Looking at a data set of a collection of cars, they are interested in exploring the relationship between a   set of variables and miles per gallon (MPG) (outcome). They are particularly interested         in the following two questions:
                Q1 “Is an automatic or manual transmission better for MPG”
                Q2 "Quantify the MPG difference between automatic and manual transmissions"
# Planned Approach -
        Experimental Design Considerations
                Simple linear comparison to:
                Multivariate incrementing by additional variable 
                Coefficients 
                Residuals
                        Influence
                        Leverage
                Correlation
                        Deviance
        Descriptive
                head
                summary
        Exploratory
                lm - simple
                lm - multivariate
                lm - nested
                lm - remove the intercept (-1)
                lm - step function 
        Inferential
                Null Hypothesis
                Alternative Hypothesis
                Power or Alpha
                Confidence Interval = .95, one or two sided?
                pValue
                
        Predictive (test)?
                ????
        Causal ~ NA
        Mechanistic ~ NA

# Software Environment
        System / session Info:
```{r, echo=TRUE}
sessionInfo()
```

# Data Processing / Cleaning
        Download, read & Date Downloaded
```{r, echo=TRUE}
'insert code'
```
        
        Raw Data - what is the available data = Motor Trend 'mtcars' data set
```{r, echo=TRUE}
head(mtcars,10)
str(mtcars)
```

# Processed Data
        Transformations / modifications / changes / etc.
        CodeBook (how you processed the data)
                Explicit steps
                        1
                        2
                        3 etc.

# Exploratory Analysis w/ Processed Data
        State the H0 & Ha hypothesis here
                Comparisons
                Causality?
                Multivariate
                Nested Analysis
                Summaries
                Boxplots
                Histograms
                Rug
                Barplot
                ABline (h/v)
                Scatterplot
                Multiple scatter plots
                Graphing - base, lattice, ggpplot2
                Heatmap
                K-Means
                Dimension Reduction
                        PCA
                        SVD
                Figures: Exploratory

# Statistical Modeling, Regression & Model Fit
        Simple Linear Regression
        Multivariate Linear Regression
        Reference Variable ??
        Coefficients / Slope
        Standard Error
        T-Vales
        pValues
        Residuals
                Leverage
                Influence
        Hatvalues
        

# Assumptions Main:
        A     
        B
        C

# Preliminary Findings: Quesions of Interest: & Interpretation of Results;
                A
                B
                C

# Inference
        Hypothesis testing
        Set Seed if necessary
        One or Two Sided Test
        Power?
        Confidence Intervals
        Standard Error
        Variance
        p-Values
        Residual Plots with diagnostics see Appendix
        Anova
        Ancova
        Hatvalues
        Influence Measures

# Conclusions / Recommendations  
        A
        B
        C
                Challenge the results ?
                Measures of uncertainty 'e'
# What are some possible alternative analyses?
        ???

# Appendix A
        Plots with Code
                Histograms
                Pairs
                QQ Plots
                Box Plots
                Residuals vs Fitted?
                Error Rates?

===
END OF DOCUMENT
===

