---
title: "Regression Models Project - Motor Trend Data 'mtcars' Miles Per Gallon Analysis"
author: "james c walmsley"
date: "12/1/2016"
output: pdf_document
---
#EXECUTIVE SUMMARY
        ## Add after completing analysis

---
# Problem Statement & Question(s) to be answered
        ## Assuming you work for Motor Trend, a magazine about the automobile industry. Looking at the data set of a collection of cars ("mtcars") you know that; 
        ## They are interested in exploring the relationship between a set of variables and the miles per gallon (mpg) outcome. 
        ## They are particularly interested in the following two questions:
                ## Q1 “Is an automatic or manual transmission better for 'mpg'”
                ## Q2 "Quantify the MPG difference between automatic and manual transmissions"
---
# Planned Approach
        Descriptive
                any(is.na)
                head(data)
                str(data)
                summary(data)
        Exploratory
                Simple linear comparisons
                Multivariate
                        Additive
                        Multiplicative
                        Missing
                        Steped
                        Coefficients 
                        Residuals
                                Influence
                                Leverage
                        Deviance
                        Anova
        Inferential
                Null Hypothesis
                Alternative Hypothesis
                Power or Alpha
                Confidence Interval = .95, one or two sided?
                pValue
                R^2
        Predictive ~ NA
        Causal ~ NA
        Mechanistic ~ NA
---
# Software Environment
        System / session Info:
```{r, echo=FALSE}
sessionInfo()
```
---
# Access the Data:
        Note on where to get this data:  
<!--         Clean the Environment -->
<!--         Download -->
<!--         Read -->
<!--         Record Date Downloaded -->
<!-- ```{r, echo=TRUE} -->
<!-- rm(list =ls()) -->
<!-- 'insert code' -->
<!-- ``` -->
---       
# Raw Data Overview: 
        Motor Trend 'mtcars' data set
```{r, echo=TRUE}
rm(list = ls())
data("mtcars")
any(is.na(mtcars))
head(mtcars,5)
```
---
# Process data:
        Transformations;  
                1 factor variables 8:11;  
                2 change variable labels in columns 8 & 9;      
                        a Note for column header 8 = vs; V = V block motor, & S = Straight block motor;  
                        b Note for column header 9 = am; A = automatic transmission = A, & M = manual transmission;  
        
```{r, echo=TRUE}
data("mtcars")
mtcars$vs <- factor(mtcars$vs, labels = c("V", "S")); mtcars$am <- factor(mtcars$am, labels = c("A", "M")); mtcars$gear <- factor(mtcars$gear); mtcars$carb <- factor(mtcars$carb)
str(mtcars)
head(mtcars)
```
---
# Exploratory Analysis
        Histograms
        Boxplots
        Rug
        Barplots
        Scatterplots
        Multiple plots
        Graphing - base, lattice, ggpplot2
        ABline (h/v)
        Confidence intervals
        Standard error
        Variance
        Fitted lines
        
        State the H0 & Ha hypothesis here
                Comparisons
                Causality?
                Multivariate
                Nested Analysis
                Summaries
                Boxplots
                Histograms
                Rug
                Barplot
                ABline (h/v)
                Scatterplot
                Multiple scatter plots
                Graphing - base, lattice, ggpplot2
                Heatmap
                K-Means
                Dimension Reduction
                        PCA
                        SVD
                Figures: Exploratory
---
# Statistical Modeling, Regression & Model Fit
        Simple Linear Regression
        Multivariate Linear Regression
                lm - simple
                lm - multivariate
                lm - nested
                lm - remove the intercept (-1)
                lm - step function 
        Coefficients / Slope
        Standard Error
        T-Vales
        pValues
        Residuals
                Leverage
                Influence
        Confidence Intervals
        Residuals
        Hatvalues
        dfbetas
        Influence Measures
        Anova
                Chisq
        Ancova
        GLM
                ?
                ?
---
# Assumptions Main:
        A     
        B
        C
---
# Preliminary Findings: Quesions of Interest: & Interpretation of Results;
                A
                B
                C
---
# Inference
        Hypothesis testing
        Set Seed, if required
        One or Two Sided Test
        Power / Alpha
        Beta = (1 - Alpha)
        Confidence Intervals (.95 one sided, .975 two sided)
        Standard Error
        Variance
        student's T-score
        Z-score
        p-Values
        Residual Plots with diagnostics see Appendix
---
# Conclusions / Recommendations  
        A
        B
        C
                Challenge the results ?
                Measures of uncertainty 'e'
---
# What are some possible alternative analyses?
        ???
---
# Appendix A
        Plots with Code
                Pairs
                Histograms
                Box Plots
                QQ Plots
                Fitted
                Residuals
                Residuals vs Fitted

===
END
===

