---
title: "Regression Models Course Project - Motor Trend Data Set - 'mtcars' Miles Per Gallon Ratings Analysis"
author: "james c walmsley"
date: "12/1/2016"
output:
  html_document: default
  pdf_document:
    fig_caption: yes
    fig_height: 5.5
---
Executive Summary:                      √
        Using a braod range of linear regressoin model variations including the step function to gauge
        best model fit we idendified a model using (wt + qsec + am) that provides an 84.966 R^2 value
        inidicating a reasonalby goo fit which practically matches a slightly different fit that was  
        devloped using the manual nested approach followed by and anove table test to check for  
        multicollinearity which also produced a reasonably good fit with an R^2 value of 85.13%.
        It must also be noted that vehicle weight is highly correlated (-86.77%) with mpg ratings and
        transmission type is relatively highly correlated with vehicle wieght at (69%).
        Problem Statement:                      √
                Backround information, problem statement & questions of interest:
                Background situation:
                        As a member of a team of data analysts for the Motor Trend Magazine we have been given  
                        a data set called "mtcars" and asked to answer some questions of interest concerning  
                        differences between automatic and  manual transmissoin types in regards to associated  
                        mpg or miles per gallon ratings within the given data set. 
                Assumptions:
                        The given data set (a sample of a larger population) for this analysis consists of (iid)  
                        independent and identically distrubuted random varialbles for 32 subjects (vehicles) of  
                        11 observations or variables.
                
                        Questions of interest for Motor Trend Magazine:
                
                        Q1 “Is an automatic or manual transmission better for 'mpg'”
                                or which type of tramsmission is associated with better mph or gas mileage ratings?
                                A1. The mean "MPG" rating of all vehicle models including both transmission types  
                                is 20.091 mpg with a 95% confint of 17.917mpg to 22.263mpg.
                        Q2 "Quantify the MPG difference between automatic and manual transmissions"
                                or assuming there is an associated difference in mpg ratings between manual and  
                                automatic type transmissions then: What is the expected difference and how accurate  
                                is this estimate based on the given data?
                                A2. The mean "MPG"  of models with automatic transmisions is 17.147 mpg, and those  
                                with manual transmisions 24.392 mpg for a difference of 7.24 mpg in favor of the  
                                manual transmission in the given data set.

Analysis Considerations:                 √
        Descriptive - any(is.na()), str() & summary(),
        Exploratory - pairsPlots(), histograms(), boxPlots(), barPlots() QQ_Plots & multiple plots
        Regression Models Analysis -
                OLS, SLR, BiVariate Regression, Multivariate Linear Regression, Heatmaps,
                HCL, PCA, SVD, Mean, T-Test, Z-Test, covariance, OLS, regression to mean (-1),
                simple linear regression, statistical linear regression, multivariable
                regression, logit & model selection, adjustments, residuals (predict fit, residual
                fit (-1)), hatvalues, variation, & dfbetas, R^2, diagnostics; ANOVA, GLMs &
                Binary GLMs, coeficients, correlation, confidence intervals, Cooks Distance,
                ChiSq-Test, VIF, binary, binomial, poisson, influence & leverage, Odds & OddRatio,
                Inferential & Predictive, Causal ~ NA, Mechanistic ~ NA
        Diagnostics
                See section on Diagnostics
        Final Model Selection strategy
                Beginning with the simmple linear regression using just one predictor (am),
                Use a bivariate model, Use a multivariate model, Use an intercept adjusted multivariate model
                Use a multivariate model removing a suspected key regressor, Use the nested multivariate process
                Use the step(function) both directions process, Usee different combinations and leave out from  
                the results of the step(function), Choose the best fit which is understandable and easy to explian
        
Technical Environment:                  √
        Environment:
                System - session Info
                Set the Working Directory
                Record the System & Session Info
                Check which packages have beeb installed

```{r, SetDirectory, echo=TRUE, eval=TRUE, include=FALSE, results='hide'}
setwd("~/Desktop/Coursera_R/7_Regression Models/RM_proj_MPG_MotorTrendData");sessionInfo();installed.packages();date()
```

Raw Data:                               √
        Clean up work space, import the data & check for missing values
        Overview: Motor Trend 'mtcars' data set:
```{r, Data_mtcars, include=FALSE, results='hide'}
rm(list=ls());library(car);library(dplyr);library(MASS);data("mtcars");any(is.na(mtcars))
```
A data frame with 32 observations on 11 variables.

[, 1]	mpg	Miles/(US) gallon
[, 2]	cyl	Number of cylinders (4,6,8)
[, 3]	disp	Displacement (cu.in.)
[, 4]	hp	Gross horsepower
[, 5]	drat	Rear axle ratio
[, 6]	wt	Weight (1000 lbs)
[, 7]	qsec	1/4 mile time
[, 8]	vs	V/S (0 = vee-block, 1 = straight-block)
[, 9]	am	Transmission (0 = automatic, 1 = manual)
[,10]	gear	Number of forward gears (3:5)
[,11]	carb	Number of carburetors (1:4,6,8)

Processed Data:                         √
        Factor columns 2 & 8:11 (cyl,vs,am,gear,carb) so there values can be used as levels
```{r, ProcessData, echo=TRUE, eval=TRUE, cache=TRUE, message=FALSE, warning=FALSE, include=FALSE, results='hide'}
data("mtcars")
for(i in c(2,8:11))mtcars[,i] <- as.factor(mtcars[,i]);
str(mtcars)
```

Descriptive Statistics:                 √

```{r, DescriptiveStats, echo=TRUE, eval=FALSE, cache=TRUE, message=FALSE, warning=FALSE, include=FALSE, results='hide'}
library(datasets);library(dplyr);data("mtcars");head(mtcars)
head(mtcars,4);mean(mtcars$mpg);sd(mtcars$mpg)
t.test(mtcars$mpg);round(t.test(mtcars$mpg)$conf.int,3)
mtcars0 <- mtcars[mtcars$am==0,];mtcars0;t.test(mtcars0$mpg)
mtcars1 <- mtcars[mtcars$am==1,];mtcars1;t.test(mtcars1$mpg)
c4 <- mtcars$mpg[mtcars$cyl==4];c4;t.test(c4,c6,var.equal = TRUE)
c6 <- mtcars$mpg[mtcars$cyl==6];c6
c8 <- mtcars$mpg[mtcars$cyl==8];c8;t.test(c6,c8,var.equal = TRUE)
```

Exploratory Analysis:                   √
        See Appendix A. Figures (pairs-plot, histogram, box-plot)
        
Statistical Modeling:                   √
        Multivarite Linear Model Finding Best Fit with Step function:
```{r ModelFitStep, echo=TRUE, eval=FALSE, message=FALSE, warning=FALSE, cache=TRUE, include=TRUE,  results='asis'}
library(stats);library(MASS)
fstp <- lm(mpg ~ ., data = mtcars)
stp <- step(fstp, trace = FALSE)
coef(summary(stp))
summary(stp)$r.squared
```

Preliminary findings: 
        # Quesions of interest: & interpretation of results:
        A Revisit the Question - Considering all regressors:
                A. Is an automatic or manual transmission better for mpg
                        The results of using multiple linear regression techniques  
                        sugggest that manual transmissions are associated with better 
                        mpg ratings than automatic transmissions
                B. Quantify the MPG difference between automatic and manual transmissions
                        On average manual transmissions provides 24.39 mpg which is 7.24 mpg
                        more than the 17.15 mpg average of the automatic transmission models
        B Primary result
                A. Are any other regressors significantly correlated with mpg rating?
                        a. model fnm6 = factor(am) + cyl + disp + hp + drat + wt
                        this model has an R^2 value of 85.13% 
                B. Further testing
                        a. using the step function in both directions selects wt, qsec and am 
                        as good predictors with an 84.96% R^2 value indicating very good 
                        predictability using this set of regresssors
                
        C Direction, Magnitude, Uncertainty
                A. 
        D Context
                A. It should be noted that vehicle weight has a strong negative correlation
                to mpg ratings and the weight of vehicle models with manual transmissions  
                range from 1.513tons to 3.570 tons and the weight of vehicles with automatic  
                transmissions range from 2.465 tons to 5.424 tons
                
        E Implications - Congruence with existing knowledge?
                A. Sedan, Sports, Luxury


Diagnostics:                            √ ? more may be needed

Hypothesis Test:                        ?
                H0 = mean(automatic transmission)mpg  =   mean(manual transmission)mpg
                Ha = mean(automatic transmission)mpg  !=  mean(manual transmission)mpg


Inference & Prediction:                 ?







Interpretation of Results:              #

Appendix A. Figures:                    √ need to review again possibly condense to one facet plot













