---
title: "Regression Models Project - Motor Trend Data 'mtcars' Miles Per Gallon Analysis"
author: "james c walmsley"
date: "12/1/2016"
output: pdf_document
---
# EXECUTIVE SUMMARY
        Information to be added on completion of analsis!





# Problem Statement / Define the Question(s)
        You work for Motor Trend, a magazine about the automobile industry. Looking at a data set of a collection of cars, they are interested in exploring the relationship between a   set of variables and miles per gallon (MPG) (outcome). They are particularly interested in the following two questions:
                Q1 “Is an automatic or manual transmission better for MPG”
                Q2 "Quantify the MPG difference between automatic and manual transmissions"
# Planned Approach -
        Experimental Design Considerations
                Simple linear comparisons
                Multivariate
                        Additive
                        Multiplicative
                        Missing
                        Steped
                Coefficients 
                Residuals
                        Influence
                        Leverage
                Correlation
                Deviance
        Descriptive
                is.na
                str
                summary
        Exploratory
                Histograms
                Boxplots
                Rug
                Barplots
                Scatterplots
                Multiple plots
                Graphing - base, lattice, ggpplot2
                ABline (h/v)
                Confidence intervals
                Standard error
                Variance
                Fitted lines
        
        Inferential
                Null Hypothesis
                Alternative Hypothesis
                Power or Alpha
                Confidence Interval = .95, one or two sided?
                pValue
                R^2
        Predictive >
        Causal ~ NA
        Mechanistic ~ NA

# Software Environment
        System / session Info:
```{r, echo=TRUE}
sessionInfo()
```

# Data Processing / Cleaning
<!--         Clean the Environment -->
<!--         Download -->
<!--         Read -->
<!--         Record Date Downloaded -->
<!-- ```{r, echo=TRUE} -->
<!-- rm(list =ls()) -->
<!-- 'insert code' -->
<!-- ``` -->
        
        Raw Data - what is the available data = Motor Trend 'mtcars' data set
```{r, echo=TRUE}
rm(list = ls())
data("mtcars")
head(mtcars,10)
str(mtcars)
head(mtcars)
```

# Processed Data
        Transformations / modifications / changes / etc.
        CodeBook (how you processed the data)
                Explicit steps
                        1 factor variables 8:11
```{r, echo=TRUE}
data("mtcars")
mtcars$vs <- factor(mtcars$vs, labels = c("V", "S")); mtcars$am <- factor(mtcars$am, labels = c("A", "M")); mtcars$gear <- factor(mtcars$gear); mtcars$carb <- factor(mtcars$carb)
str(mtcars)
head(mtcars)
```

# 1 Note for column vs; V = V block motor, & S = Straight block motor 
# 2 Note for column am; A = automatic transmission = A, & M = manual transmission


# Exploratory Analysis w/ Processed Data
        State the H0 & Ha hypothesis here
                Comparisons
                Causality?
                Multivariate
                Nested Analysis
                Summaries
                Boxplots
                Histograms
                Rug
                Barplot
                ABline (h/v)
                Scatterplot
                Multiple scatter plots
                Graphing - base, lattice, ggpplot2
                Heatmap
                K-Means
                Dimension Reduction
                        PCA
                        SVD
                Figures: Exploratory

# Statistical Modeling, Regression & Model Fit
        Simple Linear Regression
        Multivariate Linear Regression
                lm - simple
                lm - multivariate
                lm - nested
                lm - remove the intercept (-1)
                lm - step function 
        Coefficients / Slope
        Standard Error
        T-Vales
        pValues
        Residuals
                Leverage
                Influence
        Confidence Intervals
        Residuals
        Hatvalues
        dfbetas
        Influence Measures
        Anova
                Chisq
        Ancova
        GLM
                ?
                ?
# Assumptions Main:
        A     
        B
        C

# Preliminary Findings: Quesions of Interest: & Interpretation of Results;
                A
                B
                C

# Inference
        Hypothesis testing
        Set Seed, if required
        One or Two Sided Test
        Power / Alpha
        Beta = (1 - Alpha)
        Confidence Intervals (.95 one sided, .975 two sided)
        Standard Error
        Variance
        student's T-score
        Z-score
        p-Values
        Residual Plots with diagnostics see Appendix

# Conclusions / Recommendations  
        A
        B
        C
                Challenge the results ?
                Measures of uncertainty 'e'
# What are some possible alternative analyses?
        ???

# Appendix A
        Plots with Code
                Pairs
                Histograms
                Box Plots
                QQ Plots
                Fitted
                Residuals
                Residuals vs Fitted

===
END
===

