---
title: Regression Models Course Project - Motor Trend Data = 'mtcars' Miles Per
  Gallon Analysis
author: "james c walmsley"
date: "12/1/2016"
output:
  pdf_document:
    fig_caption: yes
    fig_height: 5.5
  html_document: default
---
# Executive Summary:                      
                                Interpretation of Results        
Using linear regression model variations, including the step function, to gauge model fit,  
we identified model fstp <- lm(mpg ~ (wt + qsec + am), data= mtcars) that provides an 84.966  
R^2 value indicating a reasonalby good fit and another model fit fnm6 <- lm(mpg ~ I(cyl + disp  
+ hp + drat + wt + factor(am), data = mtcars) devloped using the multivariate nested approach  
followed by and anova table test to check for multicollinearity which also produced a reasonably
good fit with an R^2 value of 85.13%. It must also be noted that vehicle weight is highly  
correlated (-86.77%) with mpg ratings and transmission type is relatively highly correlated  
with vehicle wieght at (69%).
        
                Questions of interest for Motor Trend Magazine:
                
                Q1 “Is an automatic or manual transmission better for 'mpg'”
                        or which type of transmission is associated with better mpg or gas mileage ratings?
                A1. The mean "mpg" rating of all vehicle models including both transmission types  
                is 20.09 mpg with a 95% confidence interval of 17.92 mpg to 22.26 mpg.
                                
                Q2 "Quantify the mpg difference between automatic and manual transmissions"
                What is the expected difference in mpg rating and how accurate  
                is this estimate based on the given data?
                                
                A2. The mean "mpg" of models with automatic transmisions is 17.15 mpg, with a 95%  
                confidence interval of between 14.85 mpg to 19.44 mpg and vehicles with manual  
                transmisions have a mean of 24.39 mpg for a difference of 7.24 mpg with a 95 %  
                confidence interval of between 18.49 mpg and 30.29 mpg  
                                
# Technical Environment:                  
        System - session Info; Set the Working Directory; Record the System & Session Info; Check for required packages

```{r, SetDirectory, echo=TRUE, eval=TRUE, include=FALSE, results='hide'}
setwd("~/Desktop/Coursera_R/7_Regression Models/RM_proj_MPG_MotorTrendData");sessionInfo();installed.packages();date()
```

# Raw Data:                               
        Clean up work space, import the data & check for missing values
        Overview: Motor Trend 'mtcars' data set:
```{r, Data_mtcars, include=FALSE, results='hide'}
rm(list=ls());library(car);library(dplyr);library(MASS);data("mtcars");any(is.na(mtcars))
```
A data frame with 32 observations on 11 variables.

[, 1]	mpg	Miles/(US) gallon
[, 2]	cyl	Number of cylinders (4,6,8)
[, 3]	disp	Displacement (cu.in.)
[, 4]	hp	Gross horsepower
[, 5]	drat	Rear axle ratio
[, 6]	wt	Weight (1000 lbs)
[, 7]	qsec	1/4 mile time
[, 8]	vs	V/S (0 = vee-block, 1 = straight-block)
[, 9]	am	Transmission (0 = automatic, 1 = manual)
[,10]	gear	Number of forward gears (3:5)
[,11]	carb	Number of carburetors (1:4,6,8)

# Processed Data:                         
        Factor columns 2 & 8:11 (cyl,vs,am,gear,carb) into levels
```{r, ProcessData, echo=TRUE, eval=FALSE, cache=TRUE, message=FALSE, warning=FALSE, include=FALSE, results='hide'}
data("mtcars")
for(i in c(2,8:11))mtcars[,i] <- as.factor(mtcars[,i]);
str(mtcars)
```

# Descriptive Statistics:                 

```{r, DescriptiveStats, echo=TRUE, eval=TRUE, cache=TRUE, message=FALSE, warning=FALSE, include=TRUE, results='asis'}
library(datasets);library(dplyr);data("mtcars")
head(mtcars,4);mean(mtcars$mpg);sd(mtcars$mpg)
round(t.test(mtcars$mpg)$conf.int,3)
```

# Exploratory Analysis:                   
        See Appendix A. Figures (pairs-plot, histogram, box-plot)
        
# Statistical Modeling:                   
        Multivarite Linear Model Finding Best Fit with Step function:
```{r ModelFitStep, echo=FALSE, eval=TRUE, message=FALSE, warning=FALSE, cache=TRUE, include=FALSE,  results='hide'}
library(stats);library(MASS);library(dplyr);data("mtcars")
fstp <- lm(mpg ~ ., data = mtcars)
stp <- step(fstp, trace = FALSE)
summary(stp)$r.squared
anova(stp)
```

# Multivarite Linear Model Finding Best Fit with Step function:
```{r MVLM_Step_R^2, echo=FALSE, eval=TRUE, fig.align='center', fig.height=3, fig.width=7, message=FALSE, warning=FALSE, include=TRUE, results='asis'}
library(stats);library(MASS);library(ggplot2)
fstp <- lm(mpg ~ ., data = mtcars)
stp <- step(fstp, trace = FALSE)
coef(summary(stp))
summary(stp)$r.squared
x <- (mtcars$wt + mtcars$qsec + mtcars$am)
par(mfrow = c(1, 1), mar = c(4,4,4,2))
g <- ggplot(mtcars, aes(x = x, y = mpg),)
g <- g + xlab("Model stp = (wt + qsec + am)")
g <- g + ylab("Miles per Gallon")
g <- g + geom_smooth(aes(method = "lm", shape = factor(am)))
g <- g + geom_point(aes(shape = factor(am), size=qsec, colour=wt))
g
```

# Diagnostics:                             
```{r Fig.D1_ResidFit, echo=TRUE, eval=TRUE, fig.align="center", fig.height=4, fig.width=6.5}
x <- (mtcars$wt + mtcars$qsec + as.numeric(mtcars$am))
y <- mtcars$mpg
n <- length(y)
fw <- lm(y ~ x, data = mtcars)
coef(summary(fw))
e <- resid(fw);sum(e)
#par(mfrow = c(1,1), mar = c(4,4,4,3)) # set margin
plot(x,e,
     xlab = "wt + qsec + am",
     ylab = "Residuals Fit (mpg)",
     bg = "lightblue",
     col = "black", cex = .85, pch = 21, frame = FALSE)
abline(h = 0, lwd = 2)
for(i in 1:n)
        lines(c(x[i], x[i]), c(e[i], 0), col = "red", lwd = 2)
sum(e)
```
        
# Hypothesis Test:                        
                H0 = mean(mpg[am==automatic])  =   mean(mpg[am==manual]) (REJECT)
                Ha = mean(automatic transmission)mpg  !=  mean(manual transmission)mpg (ACCEPT)
```{r, t.test_mpg, echo=TRUE, eval=TRUE,include=TRUE,results='asis'}
t.test(mpg ~ factor(am), paired = FALSE, var.equal=FALSE, data = mtcars)
```
# Inference & Prediction:                 
```{r predict, echo=TRUE, eval=FALSE, cache=TRUE, message=FALSE, warning=FALSE}
data("mtcars")
levels(mtcars$wt) <- 2
fw <- lm(mpg ~ wt, data = mtcars)
new.weights <- c(1.750, 2.125, 2.750, 3.475, 4.125, 4.700)
predict(fw, newdata = data.frame(wt = new.weights))
```

# Interpretation of Results:              
See Executive Summary

# Appendix A: Figures: Exploratory Analysis
Appendix A: Figures: Exploratory Analysis

```{r Fig.1_Pairs, echo=FALSE, eval=TRUE, fig.align='center', fig.height=3.5, fig.width=5.5, message=FALSE, warning=FALSE}

panel.cor <- function(x, y, digits = 2, prefix = "", cex.cor, ...)
{
        usr <- par("usr"); on.exit(par(usr))
        par(usr = c(0, 1, 0, 1))
        r <- abs(cor(x, y))
        txt <- format(c(r, 0.123456789), digits = digits)[1]
        txt <- paste0(prefix, txt)
        if(missing(cex.cor)) cex.cor <- 0.8/strwidth(txt)
        text(0.5, 0.5, txt, cex = cex.cor * r)
}
pairs(mtcars[c(1,6,7,9)], lower.panel=panel.smooth, upper.panel = panel.cor, cex=.85, pch=20, col="lightblue", cex.labels=2, font.lables =2, main = "MPG mtcars data")
```

```{r Fig.2_Hist, echo=FALSE, fig.align='center', fig.height=4.0, fig.width=6.5}
data(mtcars)
par(mfrow = c(2,2), mar = c(4,4,4,4)) # set margin
mtcars$vs <- factor(mtcars$vs, labels = c("V-block", "S-block")); mtcars$am <- factor(mtcars$am, labels = c("Auto", "Manual")); mtcars$gear <- factor(mtcars$gear); mtcars$carb <- factor(mtcars$carb)
hist(mtcars$mpg[mtcars$am=="Auto"], breaks=10, xlab = "Miles-Per-gallon", ylab = "Frequency", main = "Automatic Transmission", col = "yellow", xlim = range(5:40))
abline(v=mean(mtcars$mpg[mtcars$am=="Auto"]), col="red", lwd = 4)
hist(mtcars$mpg[mtcars$am=="Manual"], breaks=10, xlab = "Miles-Per-gallon", ylab = "Frequency", main = "Manual Transmission", col="lightblue", xlim = range(5:40))
abline(v=mean(mtcars$mpg[mtcars$am=="Manual"]), col="blue", lwd = 4)
##      Box Plot:
#data("mtcars");par(mfrow = c(1,1), mar = c(4,4,4,4))
mtcars$vs <- factor(mtcars$vs, labels = c("V-block", "S-block")); mtcars$am <- factor(mtcars$am, labels = c("Auto", "Manual")); mtcars$gear <- factor(mtcars$gear); mtcars$carb <- factor(mtcars$carb)
boxplot(mpg ~ am, data = mtcars, col = (c("yellow","lightblue")), ylab = "Miles-Per-Gallon", xlab = "Transmission", main =  "MPG vs Transmission Type")
```


===
        END
===














