---
title: "Regression Models Project - Motor Trend Data 'mtcars' Miles Per Gallon Analysis"
author: "james c walmsley"
date: "12/1/2016"
output: pdf_document
---
#Executive Summary
        ## Add after completing analysis

---
#Problem statement & questions to be answered
        ## Assuming you work for Motor Trend, a magazine about the automobile industry. Looking at the data set of a collection of cars ("mtcars") you know that; 
        ## They are interested in exploring the relationship between a set of variables and the miles per gallon (mpg) outcome. 
        ## They are particularly interested in the following two questions:
                ## Q1 “Is an automatic or manual transmission better for 'mpg'”
                ## Q2 "Quantify the MPG difference between automatic and manual transmissions"
---
#Analysis plan
        Descriptive
                any(is.na)
                head(data)
                str(data)
                summary(data)
        Exploratory
                Pairs
                Histograms
                Boxplots
                QQ plots
        OLS Ordinary least squares
                General least squares for linear equations
        Regression to the mean - Simple linear regression
        Statistical linear regression
                Basic - w additive Gaussian error
                Interpretation of regression coefficients (intercept, slope)
                Regression - prediction
        Residuals
                Residual variation
                Influence
                Leverage
                Estimate residual variation
                R squared
        Regression inference
        
        
        
        Multivariate regression analysis
                The linear model
                Two variable simple linear regression (additive) / (multiplicative)
                Coefficients
                Fitted values, residuals and residual variation
                Coefficients 
        Residuals
                Deviance
                Anova
        Inferential
                Null Hypothesis
                Alternative Hypothesis
                Power or Alpha
                Confidence Interval = .95, one or two sided?
                pValue
                R^2
        Predictive ~ NA
        Causal ~ NA
        Mechanistic ~ NA
---
#Software environment
        System - session Info:
```{r, echo=TRUE}
sessionInfo()
```
---
#Accessing this data
        Note on where to get this data: 
```{r, echo=TRUE}
library(datasets); data("mtcars")
head(mtcars)
```
---       
# Raw data overview: 
        Motor Trend 'mtcars' data set
```{r, echo=TRUE}
rm(list = ls())
data("mtcars")
any(is.na(mtcars))
head(mtcars,5)
```
---
# Process the data:
        Transformations;  
                1 factor variables 8:11;  
                2 change variable labels in columns 8 & 9;      
                        a Note for column header 8 = vs; V = V block motor, & S = Straight block motor;  
                        b Note for column header 9 = am; A = automatic transmission = A, & M = manual transmission;  
```{r, echo=TRUE}
data("mtcars")
mtcars$vs <- factor(mtcars$vs, labels = c("V", "S")); mtcars$am <- factor(mtcars$am, labels = c("A", "M")); mtcars$gear <- factor(mtcars$gear); mtcars$carb <- factor(mtcars$carb)
str(mtcars)
head(mtcars)
```
---
# Exploratory analysis
        Histograms
        Boxplots
        Rug
        Barplots
        Scatterplots
        Multiple plots
        Graphing - base, lattice, ggpplot2
        ABlines (h/v)
        Confidence intervals
        Standard error
        Variance
        Fitted lines
        Heatmap
        K-Means
        Dimension Reduction
                PCA
                SVD
        Figures: Exploratory see Appendix A
---
# Statistical modeling, regression & model fit
        Assumptions:
                A     
                B
                C
        Simple Linear Regression
        Multivariate Linear Regression
                lm - simple
                lm - multivariate
                lm - nested
                lm - remove the intercept (-1)
                lm - step function 
        Coefficients / Slope
        Standard Error
        T-Vales
        pValues
        Residuals
                Leverage
                Influence
        Confidence Intervals
        Residuals
        Hatvalues
        dfbetas
        Influence Measures
        Anova
                Chisq
        Ancova
        GLM
---
# Preliminary findings: quesions of interest: & interpretation of results;
        A
        B
        C
---
# Inference
        Assumptions:
                A     
                B
                C
        State H0 & Ha hypothesis'here
        Set Seed, if required
        One or Two Sided Test
        Power / Alpha
        Beta = (1 - Alpha)
        Confidence Intervals (.95 one sided, .975 two sided)
        Standard Error
        Variance
        student's T-score
        Z-score
        p-Values
        Residual Plots with diagnostics see Appendix
---
# Conclusions / recommendations  
        A
        B
        C
                1 Challenge the results ?
                2 Measures of uncertainty 'e'
---
# What are some possible alternative analyses?
        A
        B
---
# Appendix A
        Plots with Code
                Pairs
```{r, echo=FALSE}
pairs(mpg ~ ., data = mtcars)
```
                Histograms
```{r, echo=FALSE}
#hist(mtcars$mpg, breaks=15, xlab = "Miles-Per-gallon", main = "MPG Histogram", xlim = range(5:40))
data("mtcars")
require(ggplot2)
require(dplyr)
```
                Box Plots
```{r, echo=FALSE}
boxplot(mpg ~ am, data = mtcars, col = (c("grey","grey")), ylab = "Miles-Per-Gallon", xlab = "Transmission")
```
                QQ Plots
```{r, echo=FALSE}
#qqPlot(fit, main="Q-Q plot")
```
                Fitted
```{r}

```
                Residuals
```{r, echo=FALSE}
#plot(resid(fit1))
#abline(h=0)
```
                Residuals vs Fitted
```{r, echo=FALSE}


```


===
END
===

