Diagnostics:                             
        Diagnostic tests were conducted on model results in accordance with the plan for analysis  
        considerations. Besides several vehicle models exhibiting influence as denoted in the  
        Residuals vs Fitted plot such as the Chrysler Imperial, Fiat 128 & the Toyota Corrolla.  
        
        In the Normal QQ plot of Theoretical Quantiles vs Standardized residuals the Ford Pantera L  
        fell outside minus two standard deviations and the Chrysler Imperial and Fiat 128 fell  
        outside the range of positive two standard deviations of the Theoretical Quantiles.
        
        Looking at the results of the Scale-LOcation plot of the sqrt of standardized residuals vs  
        the Fitted Values three models exhibited notable results beyond 1.2 sd from center of fit.
        
        
        On the plot for cook's distance; Ford Panter L exhibits a significant cook's distance value            of between 0.5 and 1.0, in addition the Chrysler Imperial and Mercedes 230 sedan were noted  as         being somewhawt significant.

The plot of Resid(fit) produces a random distribution with a single outlier at a 
value slightly above 26, and checks out with a sum(e) of 5.995204e-15 or zero indicating 
a good fit overall.
```{r Fig.D1_ResidFit, echo=TRUE, eval=TRUE, fig.align="center", fig.height=4.5, fig.width=4.5}
x <- (mtcars$wt + mtcars$qsec + as.numeric(mtcars$am))
y <- mtcars$mpg
n <- length(y)
fw <- lm(y ~ x, data = mtcars)
coef(summary(fw))
e <- resid(fw);sum(e)
#par(mfrow = c(1,1), mar = c(4,4,4,3)) # set margin
plot(x,e,
     xlab = "wt + qsec + am",
     ylab = "Residuals Fit (mpg)",
     bg = "lightblue",
     col = "black", cex = 2.5, pch = 21, frame = FALSE)
abline(h = 0, lwd = 2)
for(i in 1:n)
        lines(c(x[i], x[i]), c(e[i], 0), col = "red", lwd = 2)
sum(e)
```



The hatvalues plot does not exhibit any obvious patterns or trends
```{r Fig.D2_hatvalues, echo=TRUE, eval=TRUE, fig.align='center', fig.height=4.5, fig.width=4.5}
par(mfrow = c(2,2), mar = c(4,4,4,3)) # set margin
x <- (mtcars$wt + mtcars$qsec + as.numeric(mtcars$am))
y <- mtcars$mpg
n <- length(y)
plot(hatvalues(lm(mpg ~ x, data = mtcars)))
plot(hatvalues(lm(mpg ~ ., data = mtcars)))
plot(hatvalues(fnm6)[order(hatvalues(fnm6), decreasing=TRUE)])
```




The dffits test shows evenly distributed random results above and
below the center at zero with a single outlier at 6 with a 31 mmpg rating
```{r Fig.D3_dft2, echo=TRUE, eval=TRUE, fig.align='center', fig.height=4.5, fig.width=4.5}
x <- (mtcars$wt + mtcars$qsec + mtcars$am)
y <- mtcars$mpg
n <- length(y)
f2 <- lm(mpg ~ x, data = mtcars)
plot(dffits(f2))
```



```{r Fig.D4_cooksDist, echo=TRUE, eval=TRUE, fig.align='center', fig.height=4.5, fig.width=4.5}
x <- (mtcars$wt + mtcars$qsec + as.numeric(mtcars$am))
y <- mtcars$mpg
n <- length(y)
plot(cooks.distance(lm(mpg ~ x, data = mtcars)))
```



```{r Fig.D5_resid(fit)/(1-hatvalues(fit)), echo=TRUE, eval=TRUE, fig.align='center',fig.height=4.5, fig.width=4.5}
data("mtcars")
x <- (mtcars$wt + mtcars$qsec + as.numeric(mtcars$am))
y <- mtcars$mpg
n <- length(y)
f2 <- lm(mpg ~ x, data = mtcars)
pressResid1 <- resid(f2)/(1-hatvalues(f2));plot(pressResid1)
```




```{r Fig.D6_dfbeta, echo=TRUE, eval=TRUE, fig.align='center', fig.height=3.5, fig.width=3.5}
x <- (mtcars$wt + mtcars$qsec + mtcars$am)
y <- mtcars$mpg
n <- length(y)
plot(dfbeta(lm(mpg ~ x, data = mtcars, family = binomial)))
```


---
