#       Simple Linear Regression Plot:
Examine the relationship between mpg ratings and transmission type; automatic vs manual
```{r Fig.4_SimpleLR, echo=FALSE, eval=FALSE, fig.align='center',fig.height=4.0, fig.width=6.5, message=FALSE, warning=FALSE}
library(dplyr);library(ggplot2)
data("mtcars");head(mtcars)
mtcars$am <- c("0" = "automatic", "1" = "manual")[as.character(mtcars$am)]
mtcars
f1 <- lm(mpg ~ factor(am), data = mtcars)
coef(summary(f1))
confint(f1)
par(mfrow = c(1,1), mar = c(2,2,2,2)) # set margin
g <- ggplot(mtcars, aes(x = factor(am), y = mpg),)
g <- g + xlab("Transmission Type")
g <- g + ylab("Miles per Gallon")
g <- g + geom_point(size = 4.0, col = "blue", alpha = .45)
g <- g + geom_abline(intercept = 9.91, slope = 7.24) # adjusted (-7.24) to fit grid
g
```


#       Bivariate Linear Model:
Examine the relationship between mpg & two variables: am & weight beacuse weight has an -86%
negative correlation with mpg (the highest single variable correlation with mpg)
```{r Fig.5a_BivariateLM, echo=FALSE, eval=FALSE, fig.align='center', fig.height=6.0, fig.width=7.5, message=FALSE, warning=FALSE, include=FALSE, results='hide'}
library(car);data("mtcars")
f2 <- lm(mpg ~ as.numeric(am) + wt, data = mtcars)
hatvalues(f2);coef(summary(f2));vif(f2);sqrt(vif(f2))
par(mfrow = c(2,2), mar = c(4,4,3,2)) # set margin
plot(f2)
```

```{r Fig.5b_BivariateLM, echo=FALSE, eval=FALSE, fig.align='center', fig.height=4, fig.width=7, message=FALSE, warning=FALSE, include=FALSE, results='hide'}
library(car);library(dplyr);library(ggplot2)
f2 <- lm(mpg ~ as.numeric(am) + wt, data = mtcars)
par(mfrow = c(1,1), mar = c(4,4,4,2))
g <- ggplot(mtcars, aes(x = as.numeric(am) + wt , y = mpg),)
g <- g + xlab("Transmission Type + Weight")
g <- g + ylab("Miles per Gallon")
g <- g + geom_smooth(aes(method = "lm", se = FALSE))
g <- g + geom_point(size = 3.5, col = "blue", alpha = .65)
g
```

#       Multivariate Linear Model (all vars):

The influence measures test indicates that the following vehicles exhibit
substatial influence on the model fit; "Merc 230","Merc 280", "Honda Civic",
"Camaro Z28","Porsche 914-2","ford Pantera L","ferrari Dino","Maserati Bora"
based on the cov.r (covariance ratio) at 2.809 or above


```{r Fig.6_MVLM(allvars), echo=FALSE, eval=c(1,2,4,7:19), fig.align='center', fig.height=7, fig.width=9, message=FALSE, warning=FALSE, include=FALSE, results='hide'}
library(car);library(ggplot2)
data("mtcars")
as.numeric(unlist(mtcars))
f3 <- lm(mpg ~ ., data = mtcars)
coef(summary(f3))
confint(f3)
vif(f3)
sqrt(vif(f3))
par(mfrow = c(1, 1), mar = c(3,3,3,2))
g <- ggplot(mtcars, aes(x = (cyl + disp + hp + drat + wt + qsec + vs + am + gear + carb), y = mpg),)
g <- g + xlab("Inclusive of All Design Components with factor of am & cyl")
g <- g + ylab("Miles per Gallon")
g <- g + geom_smooth(aes(method = "lm", se = FALSE))
g <- g + geom_point(aes(shape = factor(am), size = 4.0))
g <- g + geom_point(aes(colour = factor(cyl)))
g
```

#       Multivariate Linear Model (all vars) Adjusted:
```{r Fig.7a_MVLM(allvars,-1), echo=TRUE,  eval=TRUE, cache=TRUE, fig.align='center', fig.height=7,fig.width=9, message=FALSE, warning=FALSE, include=FALSE, results='hide'}
library(car);library(ggplot2);library(dplyr);library(MASS)
f6 <- lm(mpg ~ . -1, data = mtcars)
coef(summary(f6))
confint(f6)
par(mfrow = c(1,1), mar = c(4,4,2,2)) # set margin
g <- ggplot(mtcars, aes(x = (cyl + disp + hp + drat + wt + qsec + vs + am + gear + carb -1), y = mpg),)
g <- g + xlab("Multivariate(allvars)- Adjusted to Intercept")
g <- g + ylab("Miles per Gallon")
g <- g + geom_smooth(aes(method = "lm"))
g <- g + geom_point(aes(shape = factor(am), size = 4.0))
g <- g + geom_point(aes(colour = factor(cyl)))
g
```



#       Multivariate Linear Model (all vars - am):

Remove the am variable from the multivariate linear regression to see what
effect it has on the fit

```{r Fig.7b_MVLM(allvars,-am), echo=TRUE, eval=TRUE, fig.height=4, fig.width=4, message=FALSE, warning=FALSE, include=FALSE, results='hide'}
data("mtcars");library(ggplot2)
f4 <- lm(mpg ~ . - am, data = mtcars)
coef(summary(f4));vif(f4);sqrt(vif(f4));# plot(f3)
par(mfrow = c(1, 1), mar = c(4,4,4,2))
g <- ggplot(mtcars, aes(x = (cyl + disp + hp + drat + wt + qsec + vs + gear + carb), y = mpg),)
g <- g + xlab("All Design Components - am")
g <- g + ylab("Miles per Gallon")
g <- g + geom_smooth(aes(method = "lm", se=FALSE))
g <- g + geom_point(size = 2.5, col = "black", alpha = .85)
g <- g + geom_point(aes(shape = factor(gear), size = 4.5, colour = "white"))
g <- g + geom_point(aes(colour = factor(cyl), size = 3.5))
g <- g + geom_point(aes(size = drat), alpha = .4)
g
# par(mfrow = c(2, 2), mar = c(4,4,4,2))
# plot(f4)
```


##      Multivariate Nested LM & ANOVA table
Using a nested model approach, followed by checking for variance inflation
using the command vif we can see that cyl, disp & hp

```{r Fig.8_MVLM_ANOVA_table.test, message=FALSE, warning=FALSE, include=FALSE, results='hide'}
data("mtcars")
mtcarsO <- mtcars[order(mtcars$mpg),]
mtcarsO$am <- factor(mtcarsO$am);str(mtcarsO)
fn1 <- lm(mpg ~ factor(am), data = mtcarsO)
fnm2 <- update(fn1, mpg ~ cyl)
fnm3 <- update(fn1, mpg ~ cyl + disp)
fnm4 <- update(fn1, mpg ~ cyl + disp + hp)
fnm5 <- update(fn1, mpg ~ cyl + disp + hp + drat)
fnm6 <- update(fn1, mpg ~ cyl + disp + hp + drat + wt)
fnm7 <- update(fn1, mpg ~ cyl + disp + hp + drat + wt +  qsec) 
fnm8 <- update(fn1, mpg ~ cyl + disp + hp + drat + wt +  qsec + vs)
fnm9 <- update(fn1, mpg ~ cyl + disp + hp + drat + wt +  qsec + vs + gear)
fnm10 <- update(fn1, mpg ~ cyl + disp + hp + drat + wt +  qsec + vs + gear + carb)
anova(fn1, fnm2, fnm3, fnm4, fnm5, fnm6, fnm7, fnm8, fnm9, fnm10)
anova(fnm3,fnm6)
coef(summary(fnm6))
confint(fnm6)
summary(fnm6)$r.squared
#par(mfrow = c(2,2), mar = c(4,4,4,2))
par(mfrow = c(1, 1), mar = c(4,4,4,2))
g <- ggplot(mtcars, aes(x = (cyl + disp + hp + drat + wt), y = mpg),)
g <- g + xlab("Model fnm6 (cyl:wt + factor(am))")
g <- g + ylab("Miles per Gallon")
g <- g + geom_smooth(aes(method = "lm", se=FALSE))
g <- g + geom_point(size = 2.5, col = "black", alpha = .85)
g <- g + geom_point(aes(shape = factor(gear), size = 4.5, colour = "white"))
g <- g + geom_point(aes(colour = factor(cyl), size = 3.5))
g <- g + geom_point(aes(size = drat), alpha = .4)
g
```


#       Multivarite Linear Model Finding Best Fit with Step function:
```{r MVLM_Step_R^2, echo=TRUE, eval=TRUE, cache=TRUE, include=TRUE, results='asis'}
library(stats);library(MASS)
fstp <- lm(mpg ~ ., data = mtcars)
stp <- step(fstp, trace = FALSE)
coef(summary(stp))
summary(stp)$r.squared
par(mfrow = c(1, 1), mar = c(4,4,4,2))
g <- ggplot(mtcars, aes(x = (wt + qsec + am), y = mpg),)
g <- g + xlab("Model stp = (wt + qsec + am)")
g <- g + ylab("Miles per Gallon")
g <- g + geom_smooth(aes(method = "lm", shape = factor(am)))
g <- g + geom_point(aes(shape = factor(am), size=qsec, colour=wt))
g
```

